---
title: "ANOVA for temperature tolerance"
author: "Emma Chandler"
date: "September 11, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Libraries
```{r}
library(dplyr)
library(ggplot2)
library(rcompanion)
library(MASS)
library(moments)
library(plyr)
library(lmerTest)
library(lme4)

```



## Load Data
```{r}
remove(list = ls())

C.CMS <- read.csv("~/Horsenettle/data/CMS_cold.csv", header = TRUE)

```

## Calculate CMS 
```{r}
# Create  proportions of damage over max
trtpro <- C.CMS$exp_trt / C.CMS$exp_max
cntrpro <- C.CMS$cntr_trt / C.CMS$cntr_max

# Calculate CMS using proportions
C.CMS$CMS <- ((1 - trtpro)/(1 - cntrpro))

```




# Is there variance between regions, while accounting for block, population, and genet as random effects?
```{r}

CCMS_lmer1 <- lmer(CMS ~ location + (1|ramet) + (1|population), data = C.CMS)

(CCMS_lmer1.a <- anova(CCMS_lmer1))
summary(CCMS_lmer1.a)
(CCMS_lmer1.r <- ranova(CCMS_lmer1))



# Data Visualization
(C.CMS.p1 <- ggplot(C.CMS, aes(x = ramet, CMS, fill = location)) + 
    geom_boxplot(alpha = 0.8) +
    scale_fill_manual(values = c("#00008B", "#FF4040")) +
    theme_classic() +
    theme(axis.text.x = element_text(size = 12, angle = 0)) +
    labs(x = "Block", y = "CMS"))


```


## Is there variation in temperature tolerance between genets with ramet as replicates
```{r}
CCMSL_lmer2 <- lmer(CMS ~ ID + (1|ramet), data = C.CMS)

(CCMSL_lmer2.a <- anova(CCMSL_lmer2))
(CCMSL_lmer2.r <- ranova(CCMSL_lmer2))

unique(sort(C.CMS$ID))

#Crate summarised data
detach(package:plyr)

summary <- C.CMS %>% group_by(ID) %>% 
  summarise(mean = mean(CMS),sd = sd(CMS))

# dot plot
(C.CMS.dot <- ggplot(summary, aes(x = reorder(ID, mean), y = mean)) +
    geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd), width = 0.2) +
    geom_point(size = 3) +
    theme_classic() +
    theme(axis.text.x = element_text(size = 10, angle = 90, hjust = 0.5)) +
    labs(x = "Genotype", y = "CMS"))


```


# 1 ) Is there variance between populations?
```{r}

# Data Visualization
(C.CMS.p1 <- ggplot(C.CMS, aes(x = reorder(population, CMS), CMS, fill = location)) + 
    geom_boxplot(alpha = 0.8) +
    scale_fill_manual(values = c("#00008B", "#FF4040")) +
    theme_classic() +
    theme(axis.text.x = element_text(size = 12, angle = 0)) +
    labs(x = "Population", y = "CMS"))

```

ANOVA https://daviddalpiaz.github.io/appliedstats/analysis-of-variance.html 
https://cran.r-project.org/web/packages/lmerTest/lmerTest.pdf 
https://www.r-bloggers.com/2017/06/linear-models-anova-glms-and-mixed-effects-models-in-r/ 



# 2) Is there variance between Regions?
```{r}
# Data Visualization
(C.CMS.p2 <- ggplot(C.CMS, aes(location, CMS, fill = location)) + 
    geom_boxplot(fill = c("#00008B", "#FF4040"),alpha = 0.8, colour = c("#00008B", "#FF4040")) +
    theme_classic()+
    theme(axis.text.x = element_text(size = 12, angle = 0)) +
    labs(x = "Region", y = "CMS"))

```



# 3) Is there variation between Genotypes?
```{r}
# Data Visualization
CCMS_means <- aggregate(CMS ~ ID, data=C.CMS, mean)
CCMS_means <- CCMS_means %>% 
  mutate(location = case_when(
    grepl("CE", ID) ~ "South",
    grepl("RE", ID) ~ "South", 
    grepl("OP", ID) ~ "South",
    grepl("FR", ID) ~ "North",
    grepl("PI", ID) ~ "North"))

(C.CMS.p3 <- ggplot(CCMS_means, aes(x = reorder(ID, CMS), CMS, fill = location)) + 
    geom_col() +
    scale_fill_manual(values = c("#00008B", "#FF4040")) +
    theme_classic() +
    theme(axis.text.x = element_text(size = 12, angle = 90)) +
    labs(x = "Genotype", y = "CMS"))


```


# 4) Is there variation between ramets of a genotype?
```{r}
# Data Visualization
(C.CMS.p4 <- ggplot(C.CMS, aes(x = ramet, y = CMS, group = ID, colour = location)) + 
    geom_point() +
    geom_line() +
    scale_color_manual(values = c("#00008B", "#FF4040")) +
    theme_classic() +
    theme(axis.text.x = element_text(size = 12, angle = 0)) +
    labs(x = "Block", y = "CMS"))

```



# 5) Is there variance between location when genotype is nested in popultation and location and with block effects?
```{r}
# Data Visualization
(C.CMS.p1 <- ggplot(C.CMS, aes(x = ramet, CMS, fill = location)) + 
    geom_boxplot(alpha = 0.8) +
    scale_fill_manual(values = c("#00008B", "#FF4040")) +
    theme_classic() +
    theme(axis.text.x = element_text(size = 12, angle = 0)) +
    labs(x = "Block", y = "CMS"))
```


# Post hoc test
```{r}
# Check for homoscedasticity
par(mfrow=c(2,2))
plot(CCMS.m1)
par(mfrow=c(1,1))


# Post hoc test
(tukey.population<-TukeyHSD(CCMS.m1))

plot(tukey.population, las = 1)
```

Selecting models from aov https://www.scribbr.com/statistics/anova-in-r/ 

Using linear models and anova https://rc2e.com/linearregressionandanova 




## Check Normality
```{r}
# Original Data
hist(C.CMS$CMS, main = "Raw Data", xlab = "CMS")
shapiro.test(C.CMS$CMS)
skewness(C.CMS$CMS)

# NOT NORMALLY DISTRIBUTED!! 
# Very negatively skewed.


# cube transformation
CB_CMS<- (C.CMS$CMS)^3
hist(CB_CMS, main = "cube transformed")
shapiro.test(CB_CMS)

# square transformation
CMS<- (C.CMS$CMS)^2
hist(CMS, main = "squared transformed")
shapiro.test(CMS)

# square root transformation
SQRT_CMS <- sqrt(C.CMS$CMS)
hist(SQRT_CMS, main = "square root transformed")
shapiro.test(SQRT_CMS)

# cube root transformed
CR_CMS <- sign(C.CMS$CMS) * abs(C.CMS$CMS)^(1/3)
hist(CR_CMS, main = "cube root transformed")
shapiro.test(CR_CMS)

# log transformed
LOG_CMS <- log(C.CMS$CMS)
hist(LOG_CMS, main = "log transformed")
shapiro.test(LOG_CMS)

# reciprocal root
CMS <- -(1/(C.CMS$CMS^0.5))
hist(CMS, main = "reciprocal root transformed")
shapiro.test(CMS)

# reciprocal 
R_CMS <- -(1/(C.CMS$CMS))
hist(R_CMS, main = "reciprocal transformed")
shapiro.test(R_CMS)

# reciprocal square
RS_CMS <- -(1/(C.CMS$CMS)^2)
hist(RS_CMS, main = "reciprocal square transformed")
shapiro.test(RS_CMS)

# arcsin sqrt transformation 
AST_CMS <- asin(sqrt(C.CMS$CMS/100))
hist(AST_CMS, main = "arcsin sqrt transformed")
shapiro.test(AST_CMS)

# All histograms
par(mfrow=c(3,2))
plotNormalHistogram(C.CMS$CMS, main = "Raw")
plotNormalHistogram(CMS, main = "Square Transformed")
plotNormalHistogram(SQRT_CMS, main = "Square root Transformed")
plotNormalHistogram(CR_CMS, main = "Cuberoot Transformed")
plotNormalHistogram(LOG_CMS, main = "Log Transformed")


# The reciprocal root transformation has the most normal distribution and the highest shapiro-whilk test p-value
# Add reciprocal root transformation to data frame
C.CMS$CMS <- CMS

```

Transforming data https://daviddalpiaz.github.io/appliedstats/transformations.html
Ladder of powers http://seismo.berkeley.edu/~kirchner/eps_120/Toolkits/Toolkit_03.pdf 