---
title: "All Sporophyte"
author: "Emma Chandler"
date: "4/8/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Clear environment and set seed
```{r}
remove(list = ls())
set.seed(158358)
```

# Load Libraries
```{r message=FALSE}
library(ggplot2)
library(tidyverse)
library(reshape2)
library(ggforce)

```



# Load data
```{r}
H.CMS <- read.csv("~/Horsenettle/data/processed/H.CMS.csv", header = TRUE)
C.CMS <- read.csv("~/Horsenettle/data/processed/C.CMS.csv", header = TRUE)
H.CHPL <- read.csv("~/Horsenettle/data/processed/H.CHPL.csv", header = TRUE)
C.CHPL <- read.csv("~/Horsenettle/data/processed/C.CHPL.csv", header = TRUE)
CPS <- read.csv("~/Horsenettle/data/processed/CPS.csv", header = TRUE)
HPS <- read.csv("~/Horsenettle/data/processed/HPS.csv", header = TRUE)
```

### Rename column that are the same and convert lower to upper case
```{r}
H.CMS$HCMS <- H.CMS$CMS
C.CMS$CCMS <- C.CMS$CMS
H.CHPL$HCHPL <- H.CHPL$CHPL
C.CHPL$CCHPL <- C.CHPL$CHPL

HPS$ramet <- toupper(HPS$ramet)
CPS$ramet <- toupper(CPS$ramet)
```


### Drop columns that won't be used
```{r}

H.CMS <- H.CMS[-c(1,7,8:13)]
C.CMS <- C.CMS[-c(1,7,8:13)]
H.CHPL <- H.CHPL[-c(1,7,8:13)]
C.CHPL <- C.CHPL[-c(1,7,8:13)]
HPS <- HPS[-c(1,7)]
CPS <- CPS[-c(1,7)]
```

# Create data frame with all tolerance variables
```{r}
CMS <- merge(H.CMS, C.CMS, by = c("ID", "location", "population", "number", "ramet"), all.x = TRUE, all = TRUE)
CHPL <- merge(H.CHPL, C.CHPL, by = c("ID", "location", "population", "number", "ramet"), all.x = TRUE, all = TRUE)
PS <- merge(HPS, CPS, by = c("ID", "location", "population", "number", "ramet"), all.x = TRUE, all = TRUE)
CC <- merge(CMS, CHPL, by = c("ID", "location", "population", "number", "ramet"), all.x = TRUE, all = TRUE)
ALL <- merge(CC, PS, by = c("ID", "location", "population", "number", "ramet"), all.x = TRUE, all = TRUE)

```



# Creating a figure theme for all plots
```{r}
HStheme <- function(){            
  theme_bw()+                          
  theme(axis.text.x = element_text(size = 12), 
        text = element_text(family = "sans"), 
        axis.text.y = element_text(size = 12),
        plot.subtitle = element_text(size = 16, vjust = 1, hjust = 0.5),
        axis.title = element_text(size = 12),
        panel.grid = element_blank(),
        plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), units = , "cm"),
        plot.title = element_text(size = 20, vjust = 1, hjust = 0.5),
        legend.text = element_text(size = 12),
        legend.title = element_blank())
}
```



## All sporophytic variable box plots - Region
```{r}


All.spor_REG <- melt(subset(ALL, select=c(location, HCMS, CCMS, HCHPL, CCHPL, HPS, CPS)), id.var="location")

(All.spor.reg <- ggplot(All.spor_REG, aes(location, value, fill = location)) + 
    geom_boxplot(alpha = 0.8) +
    facet_wrap(variable ~ ., scales = "free_y", ncol = 2) +
    scale_fill_manual(values = c("#104E8B", "#CD2626")) +
    HStheme() +
    labs(x = "Region", y = "Variable Ratio"))


# Saving graph
ggsave("figures/all_spor.png", width = 7, height = 7, dpi = 300)


```



## HCMS and CCMS block effect plots
```{r}

CMS.block <- melt(subset(ALL, select=c(ramet, location, HCMS, CCMS)), id.var= c("ramet", "location"))

(CMS.block.plot <- ggplot(CMS.block, aes(x = ramet, y = value, fill = location)) + 
    geom_boxplot(alpha = 0.8) +
    facet_wrap(variable ~ ., scales = "free_y", ncol = 2) +
    scale_fill_manual(values = c("#104E8B", "#CD2626")) +
    HStheme() +
    labs(x = "Block", y = "CMS Ratio"))


# Saving graph
ggsave("figures/CMS_blockEffect.png", width = 7, height = 3, dpi = 300)

```

## Block effect for other sporophytic variables
```{r}

CHPL.block <- melt(subset(ALL, select=c(ramet, location, HCHPL, CCHPL)), id.var= c("ramet", "location"))

(CHPL.block.plot <- ggplot(CHPL.block, aes(x = ramet, y = value, fill = location)) + 
    geom_boxplot(alpha = 0.8) +
    facet_wrap(variable ~ ., scales = "free_y", ncol = 2) +
    scale_fill_manual(values = c("#104E8B", "#CD2626")) +
    HStheme() +
    labs(x = "Block", y = "CHPL Ratio"))


PS.block <- melt(subset(ALL, select=c(ramet, location, HPS, CPS)), id.var= c("ramet", "location"))

(PS.block.plot <- ggplot(PS.block, aes(x = ramet, y = value, fill = location)) + 
    geom_boxplot(alpha = 0.8) +
    facet_wrap(variable ~ ., scales = "free_y", ncol = 2) +
    scale_fill_manual(values = c("#104E8B", "#CD2626")) +
    HStheme() +
    labs(x = "Block", y = "PS Ratio"))
```



## All sporophytic variables with significant differences between genets
```{r}

# Two axes
All.spor_GEN <- melt(subset(ALL, select=c(ID, location, HCMS, CCHPL)), id.var= c("ID", "location")) %>% 
  mutate(location = as.factor(location),
         ID = reorder_within(ID, value, variable)) 

(All.spor.reg <- ggplot(All.spor_GEN, aes(x = reorder(ID, value, FUN = median, na.rm = TRUE), y = value, fill = location)) + 
    geom_boxplot(alpha = 0.8) +
    facet_wrap(variable ~ ., scales = "free_x", ncol = 1) +
    HStheme() +
    scale_x_reordered()+
    scale_fill_manual(values = c("#104E8B", "#B22222"))+
    theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.5),
          axis.text.y = element_text(size = 8))+
    labs(x = "Genet", y = "Variable Ratio"))

# Saving graph
ggsave("figures/spor_gen2.png", width = 7, height = 5, dpi = 300)


All.spor_GEN <- melt(subset(ALL, select=c(ID, location, HCMS, CCHPL)), id.var= c("ID", "location"))


# 1 Axis
(All.spor.reg <- ggplot(All.spor_GEN, aes(x = reorder(ID, value, FUN = median, na.rm = TRUE), y = value, fill = location)) + 
    geom_boxplot(alpha = 0.8) +
    facet_wrap(variable ~ ., scales = "free_y", ncol = 1) +
    HStheme() +
    scale_fill_manual(values = c("#104E8B", "#B22222"))+
    theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.5),
          axis.text.y = element_text(size = 8))+
    labs(x = "Genet", y = "Variable Ratio"))



# Saving graph
ggsave("figures/spor_gen.png", width = 7, height = 5, dpi = 300)


```

## All sporophytic variables with insignificant differences between genets
```{r message=FALSE}

# Two x-axes
IS.spor_GEN <- melt(subset(ALL, select=c(ID, location, CCMS, HCHPL, CPS, HPS)), id.var= c("ID", "location")) %>% 
  mutate(location = as.factor(location),
         ID = reorder_within(ID, value, variable))

(IS.spor.reg <- IS.spor_GEN %>% 
    group_by(location) %>% 
    ggplot(aes(reorder(ID, value, FUN = median, na.rm = TRUE), y = value, fill = location)) + 
    geom_boxplot(alpha = 0.8) +
    facet_wrap(variable ~ ., scales = "free_x", ncol = 1) +
    scale_x_reordered() +
    HStheme() +
    scale_fill_manual(values = c("#104E8B", "#B22222"))+
    theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.5),
          axis.text.y = element_text(size = 8))+
    labs(x = "Genet", y = "Variable Ratio"))

# Saving graph
ggsave("figures/ISspor_gen2.png", width = 7, height = 7, dpi = 300)


# One x-axis
IS.spor_GEN <- melt(subset(ALL, select=c(ID, location, CCMS, HCHPL, CPS, HPS)), id.var= c("ID", "location"))

IS.spor_GEN$group <- factor(IS.spor_GEN$location, c("South", "North"))

(IS.spor.reg <- IS.spor_GEN %>% 
    group_by(location) %>% 
    ggplot(aes( reorder(ID, value, FUN = median, na.rm = TRUE), y = value, fill = location)) + 
    geom_boxplot(alpha = 0.8) +
    facet_wrap(variable ~ ., scales = "free_y", ncol = 1) +
    HStheme() +
    scale_fill_manual(values = c("#104E8B", "#B22222"))+
    theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.5),
          axis.text.y = element_text(size = 8))+
    labs(x = "Genet", y = "Variable Ratio"))



# Saving graph
ggsave("figures/ISspor_gen1.png", width = 7, height = 7, dpi = 300)


```


## All sporophytic variable box plots - Population
```{r}


All.spor_POP <- melt(subset(ALL, select=c(population, location, HCMS, CCMS, HCHPL, CCHPL, HPS, CPS)), id.var= c("population","location"))

All.spor_POP$group <- factor(All.spor_POP$population, c("Cemetary", "Oil Patch", "Reserve", "Frontenac", "Prairie Island"))

(All.spor.POP <- ggplot(All.spor_POP, aes(x = group, y = value, fill = location)) + 
    geom_boxplot(alpha = 0.8, ) +
    facet_wrap(variable ~ ., scales = "free_y", ncol = 1) +
    scale_fill_manual(values = c("#104E8B", "#CD2626")) +
    HStheme() +
    labs(x = "Population", y = "Variable Ratio"))


# Saving graph
ggsave("figures/all_spor_pop.png", width = 7, height = 9, dpi = 300)




# Stacked figure showing significant pairs
my_comparisons <- list(c("Cemetary", "Oil Patch"), c("Cemetary", "Prairie Island"), c("Cemetary", "Frontenac"),   c("Cemetary", "Reserve"), c("Oil Patch", "Prairie Island"), c("Frontenac", "Oil Patch"), c("Oil Patch", "Reserve"), c("Frontenac", "Prairie Island"),   c("Prairie Island", "Reserve"), c("Frontenac", "Reserve")) # comparisons for post-hoc tests
# Edit until here


ggboxplot(All.spor_POP, x = "population", y = "value",
          fill = "location") +
  stat_compare_means(comparisons = my_comparisons, label = "p.signif") +
  scale_fill_manual(values = c("#104E8B", "#CD2626")) +
  facet_wrap(variable ~., scales = "free_y", ncol = 2) +
  HStheme() +
  theme(axis.text.x = element_text(size = 12, angle = -45, hjust = 0, vjust = 0.5)) +
  labs(x = "Population", y = "Variable Ratio")


# Saving graph
ggsave("figures/all_spor_pop_MMresults.png", width = 7, height = 10, dpi = 300)






```

