---
title: "ANOVA for temperature tolerance"
author: "Emma Chandler"
date: "September 11, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Libraries
```{r message=FALSE}

library(dplyr)
library(ggplot2)
library(lmerTest)
library(ggpubr)
library(EnvStats)
library(emmeans)
```



## Load Data
```{r}
remove(list = ls())

H.CHPL <- read.csv("~/Horsenettle/data/chlorophylldata.csv", header = TRUE)

```


## Differences between north and south for the initial measurements
```{r}

# Is there a difference between the two initial measurements
t.test(H.CHPL$treatment_initial, H.CHPL$control_initial)

# Mixed effects model checking difference between regions
HCHPL.origin <- lmer(treatment_initial ~ location + (1|ramet), data = H.CHPL)

anova(HCHPL.origin)
ranova(HCHPL.origin)

# One-way ANOVA checking difference between genets
HCHPL.origin2 <- aov(treatment_initial ~ID, H.CHPL)

summary.aov(HCHPL.origin2)


```



## Calculate CHPL Stability
```{r}

# Create columns with proportions
trtpro <- H.CHPL$treatment_final / H.CHPL$treatment_initial
cntrpro <- H.CHPL$control_final / H.CHPL$control_initial

# Calculate H.CHPL damage using proportions
H.CHPL$CHPL <- 1 - (cntrpro - trtpro)

# Export data frame for Correlation
# write.csv(x = H.CHPL, file = "~/Horsenettle/data/processed/H.CHPL.csv")

```


## Test for outliers
```{r}
# Create a boxplot of the dataset, outliers are shown as two distinct points
(outliers <- boxplot(data = H.CHPL, CHPL~location, plot=TRUE)$out)  

# Use number of potential outliers from previous code
rosner <- rosnerTest(H.CHPL$CHPL, k = 7)
rosner$all.stats
# Three Outliers- in rows 137

# Remove outliers
# H.CHPL_NoOut <- H.CHPL[-137,]
# (outliers <- boxplot(data = H.CHPL_NoOut, CHPL~location, plot=TRUE)$out)  


```



# Creating a figure theme for all plots
```{r}
HStheme <- function(){            
  theme_bw()+                          
  theme(axis.text.x = element_text(size = 12), 
        text = element_text(family = "sans"), 
        axis.text.y = element_text(size = 12),
        plot.subtitle = element_text(size = 16, vjust = 1, hjust = 0.5),
        axis.title = element_text(size = 14),
        panel.grid = element_blank(),
        plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), units = , "cm"),
        plot.title = element_text(size = 20, vjust = 1, hjust = 0.5),
        legend.text = element_text(size = 12),
        legend.title = element_blank())
}
```




# Is there variance between regions, while accounting for block, population, and genet as random effects?
```{r}
HCHPL_lmer1 <- lmer(CHPL ~ location + (1|ramet) + (1|ID), data = H.CHPL)
AIC(HCHPL_lmer1)
(HCHPL_lmer1.a <- anova(HCHPL_lmer1))
(HCHPL_lmer1.r <- ranova(HCHPL_lmer1))
summary(HCHPL_lmer1)

HCHPL_lmer1_pop <- lmer(CHPL ~ location + population + (1|ramet) + (1|ID), data = H.CHPL)
AIC(HCHPL_lmer1_pop)
(HCHPL_lmer1_pop.a <- anova(HCHPL_lmer1_pop))
(HCHPL_lmer1_pop.r <- ranova(HCHPL_lmer1_pop))
summary(HCHPL_lmer1_pop)


# Include populations as a random effect with southern populations pooled
H.CHPL <- H.CHPL %>% mutate(population_2 = case_when(location == "South" ~ "South",
                                                   population == "Prairie Island" ~ "Prairie Island",
                                                   population == "Frontenac" ~ "Frontenac"))

HCHPL_lmer1_pop <- lmer(CHPL ~ location + (1|ramet) + (1|population_2/ID), data = H.CHPL)
AIC(HCHPL_lmer1_pop)

(HCHPL_lmer1_pop.a <- anova(HCHPL_lmer1_pop))
(HCHPL_lmer1_pop.r <- ranova(HCHPL_lmer1_pop))

summary(HCHPL_lmer1_pop)

# Data Visualization
(H.CHPL.p1 <- ggplot(H.CHPL, aes(x = location, CHPL, fill = location)) + 
    geom_boxplot(alpha = 0.8) +
    scale_fill_manual(values = c("#104E8B", "#CD2626")) +
    labs(x = "Region", y = "CHPL Ratio", title = "Chlorophyll Content", subtitle = "Hot Treatment")+
    HStheme() +
    theme(legend.position = "none"))

# Saving graph
ggsave("figures/H.CHPL.p1.png", width = 7, height = 5, dpi = 300)

```

There is evidence of differences between blocks in CHPL, is there a significant difference between north and south for individual ramets in CHPL
```{r}

# Subset data from south
south <- subset(H.CHPL,  location == "South",
                 drop = TRUE)
# subset data from north
north <- subset(H.CHPL,  location == "North",
                 drop = TRUE)

# Ramet A
ramA_south <- filter(south, ramet == "A") 
ramA_north <- filter(north, ramet == "A")

ramA.ttest <- t.test(ramA_south$CHPL, ramA_north$CHPL, paired = TRUE)
ramA.ttest


# Ramet B
ramB_south <- filter(south, ramet == "B") 
ramB_north <- filter(north, ramet == "B")

ramB.ttest <- t.test(ramB_south$CHPL, ramB_north$CHPL, paired = TRUE)
ramB.ttest


# Ramet C
ramC_south <- filter(south, ramet == "C") 
ramC_north <- filter(north, ramet == "C")

ramC.ttest <- t.test(ramC_south$CHPL, ramC_north$CHPL, paired = TRUE)
ramC.ttest


# Ramet D
ramD_south <- filter(south, ramet == "D") 
ramD_north <- filter(north, ramet == "D")

ramD.ttest <- t.test(ramD_south$CHPL, ramD_north$CHPL, paired = TRUE)
ramD.ttest
```


## Is there variation in temperature tolerance between genets with ramet as replicates
```{r}
HCHPL_aov2 <- aov(CHPL ~ ID, data = H.CHPL)

summary(HCHPL_aov2)


summary <- H.CHPL %>% 
  group_by(ID, location) %>% 
  summarise(mean = mean(CHPL),sd = sd(CHPL))

# dot plot
(H.CHPL.dot <- ggplot(summary, aes(x = reorder(ID, mean), y = mean)) +
    geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd), width = 0.2) +
    geom_point(size = 3, aes(colour = location)) +
    scale_color_manual(values = c("#104E8B", "#CD2626"))+
    HStheme() +
    theme(axis.text.x = element_text(size = 10, angle = 90, hjust = 0.5)) +
    labs(x = "Genotype", y = "Chlorophyll Ratio", title = "Chlorophyll Stability", subtitle = "Hot Treatment"))

# Saving graph
ggsave("figures/H.CHPL.p2.png", width = 7, height = 5, dpi = 300)

```


## Is there variation in genets for the individual regions?
```{r}
# Is there variation in the individuals with in the independent Regions

#North
H.CHPL_N <- H.CHPL %>% 
  filter(location == "North")

HCHPL_N <- aov(CHPL ~ ID, data = H.CHPL_N)
summary(HCHPL_N)


#South
H.CHPL_S <- H.CHPL %>% 
  filter(location == "South")

HCHPL_S <- aov(CHPL ~ ID, data = H.CHPL_S)
summary(HCHPL_S)



```


# Variance between groups

```{r}

reg <- bartlett.test(CHPL ~ location, data = H.CHPL)
reg

```



# 1 ) Is there variance between populations?
```{r}

# Mixed model for populations
pop.aov <- lmer(CHPL ~ population + (1|ramet), data = H.CHPL)

(pop.aov.a <- anova(pop.aov))
(pop.aov.r <- ranova(pop.aov))


# Posthoc Test
PH.pop <-emmeans(pop.aov, list(pairwise ~ population), adjust = "tukey")
plot(PH.pop)


my_comparisons <- list(c("Cemetary", "Oil Patch"), c("Cemetary", "Prairie Island"), c("Cemetary", "Frontenac"),   c("Cemetary", "Reserve"), c("Oil Patch", "Prairie Island"), c("Frontenac", "Oil Patch"), c("Oil Patch", "Reserve"), c("Frontenac", "Prairie Island"),   c("Prairie Island", "Reserve"), c("Frontenac", "Reserve")) # comparisons for post-hoc 


ggboxplot(H.CHPL, x = "population", y = "CHPL",
          fill = "location") +
  stat_compare_means(comparisons = my_comparisons, label = "p.signif") +
  scale_fill_manual(values = c("#104E8B", "#CD2626")) +
  HStheme() +
  theme(axis.text.x = element_text(size = 12, angle = 0)) +
  labs(x = "Population", y = "CHPL")



# Saving graph
# ggsave("figures/H.CMS.POPULATION.png", width = 7, height = 5, dpi = 300)

```


