---
title: "ANOVA for temperature tolerance"
author: "Emma Chandler"
date: "September 11, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Libraries
```{r message=FALSE}

library(dplyr)
library(ggplot2)
library(lmerTest)
library(ggpubr)
library(EnvStats)
library(emmeans)
```



## Load Data
```{r}
remove(list = ls())

C.CHPL <- read.csv("~/Horsenettle/data/chlorophyll_cold.csv", header = TRUE)

```

## Calculate CHPL 
```{r}

# Create columns with proportions
trtpro <- C.CHPL$treatment_final / C.CHPL$treatment_initial
cntrpro <- C.CHPL$control_final / C.CHPL$control_initial

# Calculate C.CHPL damage using proportions
C.CHPL$CHPL <- 1 - (cntrpro - trtpro)

# Export data frame for Correlation
write.csv(x = C.CHPL, file = "~/Horsenettle/data/processed/C.CHPL.csv")

```


## Test for outliers
```{r}
# Create a boxplot of the dataset, outliers are shown as two distinct points
(outliers <- boxplot(data = C.CHPL, CHPL~location, plot=TRUE)$out)  

# Use number of potential outliers from previous code
rosner <- rosnerTest(C.CHPL$CHPL, k = 5)
rosner$all.stats
# Three Outliers- in rows 199, 201, 182

# Remove outliers
# C.CHPL_NoOut <- C.CHPL[-c(199, 201, 182),]
# (outliers <- boxplot(data = C.CHPL_NoOut, CHPL~location, plot=TRUE)$out)


```


# Creating a figure theme for all plots
```{r message=FALSE}
HStheme <- function(){            
  theme_bw()+                          
  theme(axis.text.x = element_text(size = 12), 
        text = element_text(family = "sans"), 
        axis.text.y = element_text(size = 12),
        plot.subtitle = element_text(size = 16, vjust = 1, hjust = 0.5),
        axis.title = element_text(size = 14),
        panel.grid = element_blank(),
        plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), units = , "cm"),
        plot.title = element_text(size = 20, vjust = 1, hjust = 0.5),
        legend.text = element_text(size = 12),
        legend.title = element_blank())
}
```




# Is there variance between regions, while accounting for block, population, and genet as random effects?
```{r}
CCHPL_lmer1 <- lmer(CHPL ~ location + (1|ramet) + (1|population:ID), data = C.CHPL)

(CCHPL_lmer1.a <- anova(CCHPL_lmer1))
(CCHPL_lmer1.r <- ranova(CCHPL_lmer1))



# Data Visualization
(C.CHPL.p1 <- ggplot(C.CHPL, aes(x = location, CHPL, fill = location)) + 
    geom_boxplot(alpha = 0.8) +
    scale_fill_manual(values = c("#104E8B", "#CD2626")) +
    labs(x = "Region", y = "Chlorophyll Ratio", title = "Chlorophyll Fluorescence", subtitle = "Cold Treatment")+
    HStheme() +
    theme(legend.position = "none"))

# Saving graph
ggsave("figures/C.CHPL.p1.png", width = 7, height = 5, dpi = 300)

```

There is evidence of differences between blocks in CHPL, is there a significant difference between north and south for individual ramets in CHPL
```{r}

# Subset data from south
south <- subset(C.CHPL,  location == "South",
                 drop = TRUE)
# subset data from north
north <- subset(C.CHPL,  location == "North",
                 drop = TRUE)

# Ramet A
ramA_south <- filter(south, ramet == "A") 
ramA_north <- filter(north, ramet == "A")

ramA.ttest <- t.test(ramA_south$CHPL, ramA_north$CHPL, paired = TRUE)
ramA.ttest


# Ramet B
ramB_south <- filter(south, ramet == "B") 
ramB_north <- filter(north, ramet == "B")

ramB.ttest <- t.test(ramB_south$CHPL, ramB_north$CHPL, paired = TRUE)
ramB.ttest


# Ramet C
ramC_south <- filter(south, ramet == "C") 
ramC_north <- filter(north, ramet == "C")

ramC.ttest <- t.test(ramC_south$CHPL, ramC_north$CHPL, paired = TRUE)
ramC.ttest


# Ramet D
ramD_south <- filter(south, ramet == "D") 
ramD_north <- filter(north, ramet == "D")

ramD.ttest <- t.test(ramD_south$CHPL, ramD_north$CHPL, paired = TRUE)
ramD.ttest


# Data Visualization
(C.CHPL.p1 <- ggplot(C.CHPL, aes(x = ramet, CHPL, fill = location)) + 
    geom_boxplot(alpha = 0.8) +
    scale_fill_manual(values = c("#104E8B", "#CD2626")) +
    labs(x = "Block", y = "Chlorophyll Ratio")+
    HStheme() +
    theme(legend.position = "none"))

# Saving graph
ggsave("figures/CCHPL_blocks.png", width = 7, height = 5, dpi = 300)

```

## Is there variation in temperature tolerance between genets with ramet as replicates
```{r}
CCHPL_aov2 <- aov(CHPL ~ ID, data = C.CHPL)

summary(CCHPL_aov2)


summary <- C.CHPL %>% 
  group_by(ID, location) %>% 
  summarise(mean = mean(CHPL),sd = sd(CHPL))

# dot plot
(C.CHPL.dot <- ggplot(summary, aes(x = reorder(ID, mean), y = mean)) +
    geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd), width = 0.2) +
    geom_point(size = 3, aes(colour = location)) +
    scale_color_manual(values = c("#104E8B", "#CD2626"))+
    HStheme() +
    theme(axis.text.x = element_text(size = 10, angle = 90, hjust = 0.5)) +
    labs(x = "Genotype", y = "Chlorophyll Ratio", title = "Chlorophyll Fluorescence", subtitle = "Cold Treatment"))

# Saving graph
ggsave("figures/C.CHPL.p2.png", width = 7, height = 5, dpi = 300)

```


## Is there variation in genets for the individual regions?
```{r}
# Is there variation in the individuals with in the independent Regions

#North
C.CHPL_N <- C.CHPL %>% 
  filter(location == "North")

CCHPL_N <- aov(CHPL ~ ID, data = C.CHPL_N)
summary(CCHPL_N)


#South
C.CHPL_S <- C.CHPL %>% 
  filter(location == "South")

CCHPL_S <- aov(CHPL ~ ID, data = C.CHPL_S)
summary(CCHPL_S)


```


# Variance between groups

```{r}
reg <- bartlett.test(CHPL ~ location, data = C.CHPL)
reg

```



# 1 ) Is there variation in between populations?
```{r}

# Mixed Model for populations
pop.aov <- lmer(CHPL ~ population + (1|ramet), data = C.CHPL)

(pop.aov.a <- anova(pop.aov))
(pop.aov.r <- ranova(pop.aov))

# Posthoc Test
PH.pop <-emmeans(pop.aov, list(pairwise ~ population), adjust = "tukey")
plot(PH.pop)


my_comparisons <- list(c("Cemetary", "Oil Patch"), c("Cemetary", "Prairie Island"), c("Cemetary", "Frontenac"),   c("Cemetary", "Reserve"), c("Oil Patch", "Prairie Island"), c("Frontenac", "Oil Patch"), c("Oil Patch", "Reserve"), c("Frontenac", "Prairie Island"),   c("Prairie Island", "Reserve"), c("Frontenac", "Reserve")) # comparisons for post-hoc tests
# Edit until here


ggboxplot(C.CHPL, x = "population", y = "CHPL",
          fill = "location") +
  stat_compare_means(comparisons = my_comparisons, label = "p.signif") +
  scale_fill_manual(values = c("#104E8B", "#CD2626")) +
  HStheme() +
  theme(axis.text.x = element_text(size = 12, angle = 0)) +
  labs(x = "Population", y = "CHPL")



# Saving graph
# ggsave("figures/H.CMS.POPULATION.png", width = 7, height = 5, dpi = 300)


```

